---
title: "Region of Measurement Equivalence (ROME) for Categorical Items"
date: "06/12/2021"
output: 
  html_document:
    toc: TRUE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```


```{r, include = FALSE}
library(tidyr)
library(psych)
library(lavaan)
library(coda)
library(MplusAutomation)
library(tidyverse)
library(ggplot2)
library(knitr)
theme_set(theme_classic() +
            theme(panel.grid.major.y = element_line(color = "grey92")))
```

# Read In Data

This example uses a nationally representative sample of 2002 10th graders from the Educational Longitudinal Study of 2002 (ELS: 2002), which was collected by the National Center for Education Statistics (US Department of Education, 2004).  ELS: 2002 follows student through high school and postsecondary years and includes extensive student survey items and also items from parents, teachers, and school administrators.  

```{r}
# Import dataset
els <- haven::read_sav("../../data/New ELS variable version 12.11.19.sav")
```

# Step 1: Set ROME for Group Differences

The first step is to preset the Region of Measurement Equivalence (ROME). Since there are no substantive guidelines for the cutoff values, we set the ROME following the convention suggested by Kruschke (2011):$[- 0.1s_{p}, 0.1s{p}]$. Here $s{p}$ is defined as the pooled standard deviation of the observed sum scores of the psychological test being evaluated. It can be calculated from the following equation $s_{p}=\frac{(n_{k}-1)s_{k}^{2} + (n_{j}-1)s_{j}^{2}}{n_{k}+n_{j}-2}$

```{r, set ROME}
# Select female participants
els_fe_n <- els[(els$Male == 0),]
# Select male participants
els_ma_n <- els[(els$Male == 1),]
# Compute total scores for each female participant
sum_tosc_fe_n <-rowSums(subset(els_fe_n, select = c(BYS89B, BYS89L, BYS89R, BYS89A, BYS89U)))
sum_tosc_fe_n <- describe(sum_tosc_fe_n)
# Number of female participants
n_female <- sum_tosc_fe_n$n
# Standard deviation for female group
sd_female <-sum_tosc_fe_n$sd
# Compute total scores for each male participant
sum_tosc_ma_n <-rowSums(subset(els_ma_n, select = c(BYS89B, BYS89L, BYS89R, BYS89A, BYS89U)))
sum_tosc_ma_n <-describe(sum_tosc_ma_n)
# Number of male participants
n_male <- sum_tosc_ma_n$n
# Standard deviation for male group
sd_male <-sum_tosc_ma_n$sd
# Calculate the pooled standard deviation of the observed sum scores
sd_pool <- sqrt(((n_female-1)*(sd_female^2)+(n_male-1)*(sd_male^2))/(n_female+n_male-2))
sd_pool
```

# Step 2: Obtain Parameter Posterior Distributions

## Read data from Mplus output

The second step is to identify non-invariant items. We used categorical CFA to test measurement invariance of the math self-efficacy scale across gender in Mplus, and exported the MCMC draws as a data file. Below we import the draws. Note Mplus export the burnin (first half of each chain), so we only keep the second half of each chain. 

```{r, import data from Mplus}
# Parameter
posd <- read.table("mathse_bayestest2.dat", header = FALSE)
# Be careful of the parameters, check the parameter tables from the Mplus output
draws_pre <- posd %>%
  select("chain" =1,  "iteration" = 2, "loadings"=c(3,5,7,8,9), "loadings5_g2" = 11, "factor mean"= 12, "factor variance" = 13, "threshold_g1"= 14:28, "threshold_g2" = 29:40)
table(draws_pre$chain)
# Take out burnins
draws <- draws_pre %>% 
  group_by(`chain`) %>%
  filter(row_number() > floor(n()/2))
table(draws$chain)
```

# Step 3: Compute the Expected Difference in Total Scores 

The third step is to compute the expected difference in total scores conditioned on $\eta$ and obtain its 95% HPDI.

```{r}
# latent self-efficacy 
eta <- 1
# number of categories of items
num_categories <- 4
# extract loadings from the MCMC draws
## female
load_f <- draws[, str_which(names(draws), "loadings")[1:5]]
## male
load_m <- draws[, str_which(names(draws), "loadings")[-5]]
# extract the first threshold for all items from the MCMC draws
thre1_f <- draws[, str_which(names(draws), "threshold_g1")[c(1,4,7,10,13)]]
# extract the second threshold for all items from the MCMC draws
thre2_f <- draws[, str_which(names(draws), "threshold_g1")[c(2,5,8,11,14)]]
# extract the second threshold for all items from the MCMC draws
thre3_f <- draws[, str_which(names(draws), "threshold_g1")[c(3,6,9,12,15)]]
# same for male
thre1_m <- draws[, str_which(names(draws), "threshold")[c(16,19,7,22,25)]]
thre2_m <- draws[, str_which(names(draws), "threshold")[c(17,20,8,23,26)]]
thre3_m <- draws[, str_which(names(draws), "threshold")[c(18,21,9,24,27)]]
```

```{r, compute total scores}
# functions to calculate the expected total score
tot <- function(eta, thres1, thres2, thres3, 
                loading, num_categories = 4) {
  p1 <- sapply((thres1 - loading * eta), pnorm)
  # p2 <- pnorm(thres2 - loading * eta)
  p2 <- sapply((thres2 - loading * eta), pnorm)
  p3 <- sapply((thres3 - loading * eta), pnorm)
  rowSums(num_categories - (p1 + p2 + p3))
}
# expected total scores for female
total_f <- tot(eta = eta, thres1 = thre1_f, thres2 = thre2_f,thres3 = thre3_f, loading = load_f, num_categories = 4)
describe(total_f)
# expected total scores for male
total_m <- tot(eta = eta, thres1 = thre1_m, thres2 = thre2_m,thres3 = thre3_m, loading = load_m, num_categories = 4)
# differences in expected total scores
total_diff <- total_f-total_m
describe(total_diff)
hist(total_diff)
```

```{r, obtain HPDI}
# compute HPDI for female
compute_to_fe_s <- function(eta, load = load_f, thres1 = thre1_f, thres2 = thre2_f, thres3 = thre3_f, prob = 0.95) {
  total_f <- tot(eta, thres1, thres2, thres3, loading = load)
  # Compute HPDI:
  hpdi <- coda::HPDinterval(coda::as.mcmc(total_f), prob = prob)
  # Return posterior means and HPDI
  tibble(
    eta = eta, 
    mean = mean(total_f), 
    sd = sd(total_f),
    ll = hpdi[1], 
    ul = hpdi[2]
  )
}
# test when eta =1
compute_to_fe_s(1)
# compute HPDI for male
compute_to_ma_s <- function(eta, load = load_m, thres1 = thre1_m, thres2 = thre2_m, thres3 = thre3_m, prob = 0.95) {
  total_m <- tot(eta, thres1, thres2, thres3, loading = load)
  # Compute HPDI:
  hpdi <- coda::HPDinterval(coda::as.mcmc(total_m), prob = prob)
  # Return posterior means and HPDI
  tibble(
    eta = eta, 
    mean = mean(total_m), 
    sd = sd(total_m),
    ll = hpdi[1], 
    ul = hpdi[2]
  )
}
# test when eta =1
compute_to_ma_s(1)
# compute HPDI for expected difference
compute_todiffnew_s <- function(eta, loading_f = load_f, thres1_f = thre1_f, thres2_f = thre2_f, thres3_f = thre3_f, thres1_m = thre1_m, thres2_m = thre2_m,thres3_m = thre3_m, loading_m = load_m, prob = 0.95) {
  total_f <- tot(eta = eta, thres1 = thres1_f, thres2 = thres2_f,thres3 = thres3_f, loading = loading_f)
  total_m <-  tot( eta = eta, thres1 = thres1_m, thres2 = thres2_m,thres3 = thres3_f, loading = loading_m)
  total_diff <- total_f - total_m
  # Compute HPDI:
  hpdi <- coda::HPDinterval(coda::as.mcmc(total_diff), prob = prob)
  # Return posterior means and HPDI
  tibble(
    eta = eta,
    mean = mean(total_diff), 
    sd = sd(total_diff),
    ll = hpdi[1], 
    ul = hpdi[2]
  )
}
compute_todiffnew_s(0)
compute_todiffnew_s(2)
compute_todiffnew_s(-2)
```
# Step 4: Compare the 95% HPDI with ROME and Determine Measurement Invariance

Two graphs were generated from this section. One is the expected total test scores for female and male (female score - male score), the other is the expected group differences in total test scores. Although there is a slight difference in total scores between females and males with the same trait level of math self-efficacy on this scale, it is practically negligible because the differences across the range of scores are entirely within the preset ROME [-0.412, 0.412]. 

```{r, graphs}
# Randomly draw a large number of etas
etas <- seq(-2, 2, length.out = 501)  
df_post_fe_s <- map_df(etas, compute_to_fe_s)
df_post_ma_s <- map_df(etas, compute_to_ma_s)
diff_postnew_s <- map_df(etas, compute_todiffnew_s)
# Plot the results
ggplot(df_post_fe_s, aes(x = eta, y = mean)) +
  ylab(labs(y="Expected total test scores"))+
  geom_ribbon(aes(ymin = ll, ymax = ul), alpha = .3) + 
  geom_ribbon(data = df_post_ma_s, aes(ymin = ll, ymax = ul), alpha = .3, fill = "red") + 
  geom_line(data = df_post_ma_s, aes(color = "Male"))+  
  geom_line(data = df_post_fe_s, aes(color = "Female"))+
  ggtitle("Total Test Scores for Male and Female")+
  theme(plot.title = element_text(hjust = 0.5), axis.line = element_line(color ="black", size = 0.3))+
  scale_colour_manual("Gender", values = c("Male"="red", "Female"="black"),labels = c("Female", "Male")) 
ggplot(diff_postnew_s, aes(x = eta, y = mean)) +
  geom_ribbon(aes(ymin = ll, ymax = ul), alpha = .3, fill = "grey")+
  geom_line(data = diff_postnew_s, col = "orange")+
  geom_hline(yintercept = 0.412, col = "red", linetype = "dashed")+
  geom_hline(yintercept = -0.412, col = "red", linetype = "dashed")+
  ggtitle("Expected Group Differences in Total Test Scores")+
  labs(y="Group Difference")+
  theme(plot.title = element_text(hjust = 0.5), axis.line = element_line(color ="black", size = 0.3))
```


