---
title:  "Region of Measurement Equivalence (ROME) for Continuous Indicators"
date: "06/15/2021"
output: 
  html_document:
    toc: TRUE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(lavaan)
library(psych)
library(ggplot2)
library(blavaan)
library(knitr)
theme_set(theme_classic() +
            theme(panel.grid.major.y = element_line(color = "grey92")))
```

# Read In Data

This example uses a nationally representative sample of 2002 10th graders from the Educational Longitudinal Study of 2002 (ELS: 2002), which was collected by the National Center for Education Statistics (US Department of Education, 2004).  ELS: 2002 follows student through high school and postsecondary years and includes extensive student survey items and also items from parents, teachers, and school administrators.  

```{r}
# Import dataset
els <- haven::read_sav("../../data/New ELS variable version 12.11.19.sav")
```

# Fit the Frequentist Model

In this part, conventional frequentist method was used to detect partial invariance. In this section, a single factor multigroup CFA model with five indicators was fit to data. As discussed in Schaefer (2009) and suggested by the modification indices, we freed two pairs of item covariances: item 1 (BYS89A) and item 2 (BYS89B), item2 (BYS89B) and item3 (BYS89L).

```{r}
# parallel analysis
pa.els <- els%>%
  select(BYS89A, BYS89B, BYS89L, BYS89R, BYS89U)%>%
  fa.parallel(cor ="poly")
# percentage of variance accounted for by each factor
100*pa.els$fa.values/length(pa.els$fa.values)
# Preliminary model
model_o <-  'Math =~ BYS89A + BYS89B + BYS89L + BYS89R + BYS89U'
# Convert variable "Male" to a factor
els$Male_n <- as_factor(els$Male)
# Fit the configural invariant model
fit_o <- cfa(model_o,data =els, group ="Male_n", missing ='fiml', std.lv = TRUE)
# check model fit
#summary(fit_o, fit.measures = TRUE)
modificationindices(fit_o, sort. = TRUE)
# Define the model
Model <- 'Math =~ BYS89A + BYS89B + BYS89L + BYS89R + BYS89U
          BYS89A ~~ BYS89B
          BYS89B ~~ BYS89L' 
# Number of indicators
n_ind <- 5
fit <- cfa(Model,data =els, missing ='fiml', std.lv = TRUE)
# summary(fit, fit.measures = TRUE)
```

## Find Non-invariant Items

The second section focuses on finding non-invariant items. First, both the configural invariance model and weak invariance model were fit to data. The Chi-square difference test was significant,indicating some items were non-invariaant. 

```{r}
fit_cor_confi <- cfa(Model, data = els, group = "Male_n", missing='fiml', std.lv= TRUE)
# summary(fit_cor_confi, fit.measures = TRUE)
# modificationIndices(fit_cor_confi, sort. = TRUE)
# Fit the weak invariant model
fit_cor_weak <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", missing='fiml', std.lv= TRUE)
# Get model fit information
#summary(fit_cor_weak, fit.measures = TRUE)
# Chi-square difference test
lavTestLRT(fit_cor_confi, fit_cor_weak)
```

After conducting a series of Chi-square difference tests, item 5 (BYS89U) was found to be non-invariant.

```{r}
# fit partial weak invariance model
fit_cor_weak_1 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89A"), missing = 'fiml', std.lv = TRUE)
fit_cor_weak_2 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89B"),  missing = 'fiml', std.lv= TRUE)
fit_cor_weak_3 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89L"),  missing = 'fiml', std.lv= TRUE)
fit_cor_weak_4 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89R"),  missing = 'fiml',std.lv= TRUE)
fit_cor_weak_5 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89U"),  missing='fiml', std.lv= TRUE)
# chi-square difference test
lavTestLRT(fit_cor_weak, fit_cor_weak_1)
lavTestLRT(fit_cor_weak, fit_cor_weak_2)
lavTestLRT(fit_cor_weak, fit_cor_weak_3)
lavTestLRT(fit_cor_weak, fit_cor_weak_4)
lavTestLRT(fit_cor_weak, fit_cor_weak_5)
```

```{r, results = 'hide'}
# item 5 is noninvariant, continue searching 
fit_cor_weak_51 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89U", "Math =~ BYS89A"),  missing='fiml', std.lv= TRUE)
fit_cor_weak_52 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89U", "Math =~ BYS89B"),  missing='fiml', std.lv= TRUE)
fit_cor_weak_53 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89U", "Math =~ BYS89L"),  missing='fiml', std.lv= TRUE)
fit_cor_weak_54 <- cfa(Model, data = els, group = "Male_n", group.equal = "loadings", group.partial = c("Math =~ BYS89U", "Math =~ BYS89R"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_weak_5, fit_cor_weak_51)
lavTestLRT(fit_cor_weak_5, fit_cor_weak_52)
lavTestLRT(fit_cor_weak_5, fit_cor_weak_53)
lavTestLRT(fit_cor_weak_5, fit_cor_weak_54)
# get model summary
# summary(fit_cor_weak_5, fit.measures = TRUE)
```

The conventional measurement invariance test continued, suggesting item 2(BYS89B), 4 (BYS89R) and 5 (BYS89U) have noninvariant intercepts across gender.

```{r}
# Fit the strong invariant model
fit_cor_strong <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_strong, fit_cor_weak_5)
```

```{r}
# Fit partial strong invariance models
# Item2 is noninvariant
fit_cor_strong_5i1 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89A ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i2 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i3 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89L ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i4<- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89R ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i5<- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89U ~ 1"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_strong_5i1,fit_cor_strong)
lavTestLRT(fit_cor_strong_5i2,fit_cor_strong)
lavTestLRT(fit_cor_strong_5i3,fit_cor_strong)
lavTestLRT(fit_cor_strong_5i4,fit_cor_strong)
lavTestLRT(fit_cor_strong_5i5,fit_cor_strong)
# summary(fit_cor_strong_5i2, fit.measures = TRUE)
```

```{r, results = 'hide'}
# Item 4 is noninvariant
fit_cor_strong_5i21 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89A ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i23 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89L ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i24 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i25 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89U ~ 1"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_strong_5i21,fit_cor_strong_5i2)
lavTestLRT(fit_cor_strong_5i23,fit_cor_strong_5i2)
lavTestLRT(fit_cor_strong_5i24,fit_cor_strong_5i2)
lavTestLRT(fit_cor_strong_5i25,fit_cor_strong_5i2)
```

```{r, results = 'hide'}
# Item 5 is noninvariant
fit_cor_strong_5i241 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89A ~1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i243 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89L ~1"),  missing='fiml', std.lv= TRUE)
fit_cor_strong_5i245 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~1"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_strong_5i241,fit_cor_strong_5i24)
lavTestLRT(fit_cor_strong_5i243,fit_cor_strong_5i24)
lavTestLRT(fit_cor_strong_5i245,fit_cor_strong_5i24)
# summary(fit_cor_strong_5i245, fit.measures = TRUE)
```

We further move to test unique factor variances and covariances. Results show items 4 and 5 have noninvariant unique factor variances. The covariance between item 1 and item 2 is noninvariant across gender.

```{r, results = 'hide'}
# Fit the partial strict invariant model (fix unique factor variances)
fit_cor_stric_5i245 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_stric_5i245,fit_cor_strong_5i245)
# searching for noninvariant unique factor variances
# item 5
fit_cor_stric_5i245u1 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1", "BYS89A ~~ BYS89A"),  missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u2 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1", "BYS89B ~~ BYS89B"),  missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u3 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1", "BYS89L ~~ BYS89L"),  missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u4 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1", "BYS89R ~~ BYS89R"),  missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u5 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U"),  missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_stric_5i245,fit_cor_stric_5i245u1)
lavTestLRT(fit_cor_stric_5i245,fit_cor_stric_5i245u2)
lavTestLRT(fit_cor_stric_5i245,fit_cor_stric_5i245u3)
lavTestLRT(fit_cor_stric_5i245,fit_cor_stric_5i245u4)
lavTestLRT(fit_cor_stric_5i245,fit_cor_stric_5i245u5)
```

```{r, results ='hide'}
#item 4 is noninvariant
fit_cor_stric_5i245u51 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89A ~~ BYS89A"), missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u52 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89B ~~ BYS89B"), missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u53 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89L ~~ BYS89L"), missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u54 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R"), missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_stric_5i245u51,fit_cor_stric_5i245u5)
lavTestLRT(fit_cor_stric_5i245u52,fit_cor_stric_5i245u5)
lavTestLRT(fit_cor_stric_5i245u53,fit_cor_stric_5i245u5)
lavTestLRT(fit_cor_stric_5i245u54,fit_cor_stric_5i245u5)
```

```{r, results = 'hide'}
# continue searching
fit_cor_stric_5i245u541 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89A ~~ BYS89A"), missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u542 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89B ~~ BYS89B"), missing='fiml', std.lv= TRUE)
fit_cor_stric_5i245u543 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89L ~~ BYS89L"), missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_stric_5i245u541,fit_cor_stric_5i245u54)
lavTestLRT(fit_cor_stric_5i245u542,fit_cor_stric_5i245u54)
lavTestLRT(fit_cor_stric_5i245u543,fit_cor_stric_5i245u54)
```


```{r, results = 'hide'}
# Fit the partial strict invariant model (fix unique factor covariances)
fit_cor_striccov_5i245u54 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals", "residual.covariances"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R"), missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_striccov_5i245u54, fit_cor_stric_5i245u54)
```

```{r, results='hide'}
fit_cor_striccov_5i245u54c1 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals", "residual.covariances"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89A ~~ BYS89B"), missing='fiml', std.lv= TRUE)
fit_cor_striccov_5i245u54c2 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals", "residual.covariances"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89B ~~ BYS89L"), missing='fiml', std.lv= TRUE)
lavTestLRT(fit_cor_striccov_5i245u54, fit_cor_striccov_5i245u54c1)
lavTestLRT(fit_cor_striccov_5i245u54, fit_cor_striccov_5i245u54c2)
```

To conclude, item 1 (BYS89A) and 3 (BYS89L) are strict invariant, item 2 (BYS89B) and 4 (BYS89R) are weak invariant, and item 5 is non-invariant. 

```{r}
# final frequentist cfa model
fit_cor_striccov_5i245u54c1 <- cfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals", "residual.covariances"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89A ~~ BYS89B"), missing='fiml', std.lv= TRUE)
summary(fit_cor_striccov_5i245u54c1, fit.measures = TRUE)
```

# Step 1: Set ROME for Group Differences

In this section, we preset the Region of Measurement Equivalence (ROME). Since there are no substantive guidelines for the cutoff values, we set the ROME following the convention suggested by Kruschke (2010): $[- 0.1s_{p}, 0.1s{p}]$. Here $s{p}]$ is defined as the pooled standard deviation of the observed sum scores of the psychological test being evaluated. It can be calculated from the following equation $s_{p}=\frac{(n_{k}-1)s_{k}^{2} + (n_{j}-1)s_{j}^{2}}{n_{k}+n_{j}-2}$

```{r}
# Select female participants
els_fe_n <- els[(els$Male == 0),]
# Select male participants
els_ma_n <- els[(els$Male == 1),]
# Compute total scores for each female participant
sum_tosc_fe_n <-rowSums(subset(els_fe_n, select = c(BYS89B, BYS89L, BYS89R, BYS89A, BYS89U)))
sum_tosc_fe_n <- describe(sum_tosc_fe_n)
# Number of female participants
n_female <- sum_tosc_fe_n$n
# Standard deviation for female group
sd_female <-sum_tosc_fe_n$sd
# Compute total scores for each male participant
sum_tosc_ma_n <-rowSums(subset(els_ma_n, select = c(BYS89B, BYS89L, BYS89R, BYS89A, BYS89U)))
sum_tosc_ma_n <-describe(sum_tosc_ma_n)
# Number of male participants
n_male <- sum_tosc_ma_n$n
# Standard deviation for male group
sd_male <-sum_tosc_ma_n$sd
# Calculate the pooled standard deviation of the observed sum scores
sd_pool <- sqrt(((n_female-1)*(sd_female^2)+(n_male-1)*(sd_male^2))/(n_female+n_male-2))
sd_pool
```

# Step 2: Obtain Parameter Posterior Distributions

## Fit the Bayesian MG-CFA Model

After partial invariance was detected, a Bayesian MG-CFA model was fit to data.

```{r}
# Fit the Bayesian MG-CFA model
# bcfa_mod <-bcfa(Model, data = els, group = "Male_n", group.equal = c("loadings", "intercepts","residuals", "residual.covariances"), group.partial = c("Math =~ BYS89U", "BYS89B ~ 1", "BYS89R ~ 1", "BYS89U ~ 1","BYS89U ~~ BYS89U", "BYS89R ~~ BYS89R", "BYS89A ~~ BYS89B"), bcontrol = list(cores = 3), std.lv = TRUE)
# # # Save the results to file
#  saveRDS(bcfa_mod, "blava_res_ml.RDS")
# Read the results from file
blava_res <- read_rds("blava_res_ml.RDS")
# Summary of the model
summary(blava_res, fit.measures = TRUE)
# MCMC chain output 
blava_draws <- blavInspect(blava_res, "draws")
# Combine results from three chains to one chain
blava_draw_sam <- map_df(blava_draws, as.data.frame)
```

# Step 3: Compute the Expected Difference in Total Scores 

In this part, we compute the expected difference in total scores conditioned on $\eta$ and obtain its 95% HPDI.

## Sum of Loadings for Each Group

First, the sum of loadings for each group, as well as the group difference in sum of loadings were calculated.

```{r}
# sum of loadings for female group by indices
# sum_ld_fe_s <- 1 + blava_draw_sam[,1]+blava_draw_sam[,2]+blava_draw_sam[,3]+blava_draw_sam[,4]
# hist(sum_ld_fe_s)

# sum of loadings for female group by str_which function
sum_ld_f <- rowSums(blava_draw_sam[ , str_which(names(blava_draw_sam), "ly")[1:n_ind]])
hist(sum_ld_f)
# sum of loadings for male group by str_which function
sum_ld_m <- rowSums(blava_draw_sam[ , str_which(names(blava_draw_sam), "ly")[(n_ind + 1):(2*n_ind)]])
hist(sum_ld_m)
# Group difference in sum of factor loadings
diff_sum_ldnew_s <- sum_ld_f-sum_ld_m
hist(diff_sum_ldnew_s)
```

## Sum of Intercepts for Each Group

Second, the sum of intercepts for each group, as well as the group difference in sum of intercepts were calculated.

```{r}
# sum of intercepts for female group 
sum_in_f <- rowSums(blava_draw_sam[ , str_which(names(blava_draw_sam), "Nu")[1:(n_ind)]])
hist(sum_in_f)
# sum of intercepts for male group
sum_in_m <- rowSums(blava_draw_sam[ , str_which(names(blava_draw_sam), "Nu")[(n_ind+1):(n_ind+5)]])
hist(sum_in_m)
# Group difference in sum of intercepts
diff_sum_innew_s <- sum_in_f - sum_in_m
hist(diff_sum_innew_s, xlim = c(-0.412, 0.412))
```

## Group Difference in Total Test Scores and 95% HPDI

Third, the expected total test scores for each group and the expected group difference in total test scores were calculated.

```{r}
# Define function to compute the 95% HPDI for Female
compute_to_fe_s <- function(eta, si = sum_in_f, sl = sum_ld_f,
                        prob = 0.95) {
  tot_sam10 <- sl * eta + si
  # Compute HPDI:
  hpdi <- coda::HPDinterval(coda::as.mcmc(tot_sam10), prob = prob)
  # Return posterior means and HPDI
  tibble(
    eta = eta, 
    mean = mean(tot_sam10), 
    sd = sd(tot_sam10),
    ll = hpdi[1], 
    ul = hpdi[2]
  )
}
compute_to_fe_s(0)
#Define function to compute the 95% HPDI for Male
compute_to_ma_s <- function(eta, si = sum_in_m, sl = sum_ld_m,
                        prob = 0.95) {
  tot_sam12 <- sl * eta + si
  # Compute HPDI:
  hpdi <- coda::HPDinterval(coda::as.mcmc(tot_sam12), prob = prob)
  # Return posterior means and HPDI
  tibble(
    eta = eta, 
    mean = mean(tot_sam12), 
    sd = sd(tot_sam12),
    ll = hpdi[1], 
    ul = hpdi[2]
  )
}
compute_to_ma_s(0)
#Define function to compute the 95% HPDI for group difference
compute_todiffnew_s <- function(eta, si = diff_sum_innew_s, sl = diff_sum_ldnew_s, prob = 0.95) {
    tot_samdiff <- sl * eta + si
  # Compute HPDI:
  hpdi <- coda::HPDinterval(coda::as.mcmc(tot_samdiff), prob = prob)
  # Return posterior means and HPDI
  tibble(
    eta = eta,
    mean = mean(tot_samdiff), 
    sd = sd(tot_samdiff),
    ll = hpdi[1], 
    ul = hpdi[2]
  )
}
compute_todiffnew_s(0)
compute_todiffnew_s(2)
compute_todiffnew_s(-2)
```

# Step 4: Compare the 95% HPDI with ROME and Determine Measurement Invariance

Two graphs were generated from this section. One is the expected total test scores for female and male (female score - male score), the other is the expected group differences in total test scores. Although there is a slight difference in total scores between females and males with the same trait level of math self-efficacy on this scale, it is practically negligible because the differences across the range of scores are entirely within the preset ROME [-0.412, 0.412]. 

```{r}
# Randomly draw a large number of etas
etas <- seq(-2, 2, length.out = 501)  
df_post_fe_s <- map_df(etas, compute_to_fe_s)
df_post_ma_s <- map_df(etas, compute_to_ma_s)
diff_postnew_s <- map_df(etas, compute_todiffnew_s)
# Plot the results
ggplot(df_post_fe_s, aes(x = eta, y = mean)) +
  ylab(labs(y="Expected total test scores"))+
  geom_ribbon(aes(ymin = ll, ymax = ul), alpha = .3) + 
  geom_ribbon(data = df_post_ma_s, aes(ymin = ll, ymax = ul), alpha = .3, fill = "red") + 
  geom_line(data = df_post_ma_s, aes(color = "Male"))+  
  geom_line(data = df_post_fe_s, aes(color = "Female"))+
  ggtitle("Total Test Scores for Male and Female")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_colour_manual("Gender", values = c("Male"="red", "Female"="black"),labels = c("Female", "Male")) 
ggplot(diff_postnew_s, aes(x = eta, y = mean)) +
  geom_ribbon(aes(ymin = ll, ymax = ul), alpha = .3, fill = "grey")+
  geom_line(data = diff_postnew_s, col = "orange")+
  geom_hline(yintercept = 0.412)+
  geom_hline(yintercept = -0.412)+
  ggtitle("Expected Group Differences in Total Test Scores")+
  labs(y="Group Difference")+
  theme(plot.title = element_text(hjust = 0.5))
```
